\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{booktabs}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}

\lstset{
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue},
  commentstyle=\color{gray},
  breaklines=true,
  frame=single,
  xleftmargin=1em,
}

\title{Machine-Checked Proofs of the $m=4$ and $m=6$ \\
Directed Hamiltonian Torus Decompositions}

\author{
  Kevin O'Connor\\
  \texttt{default.main.address@gmail.com}
}

\date{March 2026}

\begin{document}

\maketitle

\begin{abstract}
We give machine-checked proofs of the $m=4$ and $m=6$ cases of the Directed
Hamiltonian Torus Decomposition problem, verified by the Lean~4 kernel via
the \texttt{decide} tactic. A general construction discovered via
autonomous LLM exploration covers all even $m \ge 8$, but the cramped
topological spaces of the small even cases fall outside its scope.
We find valid decompositions for the $4 \times 4 \times 4$ torus
(64 vertices) and the $6 \times 6 \times 6$ torus (216 vertices, 648 arcs)
using Simulated Annealing with a non-monotonic cooling schedule. The $m=6$
witness was found in 100 seconds on an Apple~M4 processor. Both witnesses
are verified inside Lean's trusted kernel with zero unproven subgoals and
no external dependencies.
\end{abstract}

\section{Introduction}

Consider a directed graph $G_m$ with vertex set
$V = \{(i, j, k) : 0 \le i, j, k < m\}$ and directed edges from each vertex
$(i,j,k)$ to $(i{+}1, j, k)$, $(i, j{+}1, k)$, and $(i, j, k{+}1)$, all
modulo~$m$. This graph is isomorphic to the Cartesian product of three
directed cycles $C_m \square C_m \square C_m$; the broader study of
Hamiltonian decompositions in such products has been a sustained focus
in combinatorics~\cite{darijani2025}. The graph has $m^3$ vertices and
$3m^3$ arcs.

\begin{definition}
A \emph{Directed Hamiltonian Torus Decomposition} of $G_m$ is a partition of
the $3m^3$ arcs into exactly three directed Hamiltonian cycles, each of
length~$m^3$.
\end{definition}

Knuth posed this problem in Volume~4B of \emph{The Art of Computer
Programming}~\cite{knuth2026}. In February 2026, Claude Opus~4.6, guided
by Filip Stappers, discovered a construction for all odd~$m > 1$ through
31 autonomous exploration steps; Knuth subsequently published a formal
mathematical proof~\cite{knuth2026}. During that exploration, Claude also
found isolated solutions for $m = 4$, $m = 6$, and $m = 8$, but could
not generalize them. Shortly after, under the direction of Ho Boon
Suan, GPT-5.4~Pro autonomously generated a closed-form algorithm for
all even $m \ge 8$~\cite{hoboonsuan2026}, using a splice gadget on
the final two modular fibers that requires $m \ge 8$ for topological
stability. This construction was further explored by
Aquino-Michaels~\cite{aquino2026}.
The cases $m = 4$ and $m = 6$ fall outside that construction.

Morrison formalized Knuth's odd-$m$ proof in Lean~4~\cite{morrison2026},
employing a similar \texttt{decide}-based kernel reflection strategy.
To our knowledge, no machine-checked proof of any even-$m$ case has been
published.

We make two contributions:
\begin{enumerate}
  \item Machine-checked proofs of the $m = 4$ and $m = 6$ decompositions,
  stated as \texttt{Prop}-valued theorems in Lean~4 and verified by
  \texttt{decide} inside the trusted kernel.
  \item A stochastic search that finds the $m=6$ witness in the
  $6^{215}$-element search space, using a corrected energy function and a
  non-monotonic cooling schedule with the same hyperparameters as $m=4$.
\end{enumerate}

\section{Method}

\subsection{State Space}

At each vertex, the three outgoing arcs are assigned to three colors
$\{0, 1, 2\}$. There are $3! = 6$ valid assignments per vertex, giving a
full state space of $6^{m^3}$.

We identify the vertex set $V$ with the index set $\{0, \ldots, m^3{-}1\}$
via the lexicographic bijection $(i,j,k) \mapsto im^2 + jm + k$.
A state is an array $\sigma \in \{0, \ldots, 5\}^{m^3}$, where
$\sigma[v]$ selects which permutation of $\{0,1,2\}$ is applied to the
outgoing directions $(+x, +y, +z)$ at vertex~$v$. The value
$\sigma[v] \in \{0,\ldots,5\}$ encodes one of the six permutations
of~$S_3$ as shown in Table~\ref{tab:permutations}.

\begin{table}[h]
\centering
\caption{Permutation mapping from state value to spatial direction.}
\label{tab:permutations}
\begin{tabular}{@{}cccc@{}}
\toprule
$\sigma[v]$ & Color 0 & Color 1 & Color 2 \\
\midrule
0 & $+x$ & $+y$ & $+z$ \\
1 & $+x$ & $+z$ & $+y$ \\
2 & $+y$ & $+x$ & $+z$ \\
3 & $+z$ & $+x$ & $+y$ \\
4 & $+y$ & $+z$ & $+x$ \\
5 & $+z$ & $+y$ & $+x$ \\
\bottomrule
\end{tabular}
\end{table}

The origin $(0,0,0)$ corresponds to index~0. We fix $\sigma[0] = 0$
without loss of generality, breaking the color symmetry and reducing the
search space to $6^{m^3-1}$: this is $6^{63} \approx 10^{49}$ for
$m=4$ and $6^{215} \approx 10^{167}$ for $m=6$.

\subsection{Energy Function}

Each color subgraph has out-degree~1 at every vertex by construction, but
the in-degree is not guaranteed to be~1. A vertex may receive zero or
multiple incoming arcs of a given color from its three axial predecessors.
A graph with out-degree~1 but variable in-degree is a functional graph:
its components are directed cycles with trees feeding into them. A naive
energy function that counts only the number of cycles can assign $E = 0$
to such a non-Hamiltonian topology.

\begin{definition}
Let $C_c(\sigma)$ be the number of weakly connected components in the
color-$c$ functional graph, and let $D_c(\sigma)$ be the number of vertices
with in-degree $\ne 1$ for color~$c$. The energy is
\[
  E(\sigma) = \sum_{c=0}^{2} \bigl(C_c(\sigma) - 1\bigr)
            + \sum_{c=0}^{2} D_c(\sigma).
\]
\end{definition}

Because every vertex has out-degree exactly~1, each weakly connected
component contains exactly one directed cycle. $E = 0$ holds if and only
if each color subgraph is a single cycle that visits every vertex. Computing $E$ from scratch takes $O(m^3)$ time. For $m=6$,
we also implemented an incremental evaluator: each single-vertex mutation
updates the in-degree counts in $O(1)$ and recounts cycles only for the
affected colors. This was validated by a 300{,}000-step fuzz test asserting
exact agreement with full recalculation at every mutation.

\subsection{Cooling Schedule}

We use the Metropolis-Hastings algorithm with a single-vertex mutation
operator: a random vertex $v \ne 0$ has its permutation changed to one of
the five alternatives.

The cooling schedule is $T_k = T_0 \cdot \alpha^k$ with $T_0 = 50$ and
$\alpha = 0.999998$. When the best energy has not improved for $W$
iterations, the temperature is reset to
$T_r = \max(1, E_{\text{best}}^{2/5})$ and the stagnation window doubles
($W \leftarrow 2W$). Upon improvement, $W$ resets to its initial value.
This non-monotonic schedule allows the search to escape local
minima~\cite{kirkpatrick1983,marinari1992}.
The exponent $2/5$ in the reheat formula was selected empirically
via hyperparameter search; its effectiveness likely reflects the
sub-linear coupling between basin depth and the temperature required
to escape it in sparse functional graph topologies.

The same cooling rate, reheat formula, and initial temperature were used
for both $m=4$ and $m=6$ without modification.

\section{Results}

\begin{table}[h]
\centering
\caption{Simulated Annealing results.}
\label{tab:results}
\begin{tabular}{@{}llcccl@{}}
\toprule
Graph & Engine & $\alpha$ & Reheat & Best $E$ & Time \\
\midrule
$G_4$ & Baseline & 0.999998 & $E^{2/5}$, backoff & 0 & 93\,s \\
$G_4$ & Incremental & 0.999998 & $E^{2/5}$, backoff & 0 & 3.1\,s \\
$G_6$ & Incremental & 0.999998 & $E^{2/5}$, backoff & 0 & 99.5\,s \\
\bottomrule
\end{tabular}
\end{table}

The baseline engine computes $E$ from scratch on every mutation. The
incremental engine updates $E$ in place and rolls back rejected mutations,
giving roughly 12$\times$ higher throughput on $m=6$. All three runs used
identical hyperparameters.

The $m=4$ baseline solution was found at restart~12, iteration~1{,}366{,}834.
With the incremental engine, $m=4$ was solved at restart~3 in 3.1 seconds.
The $m=6$ solution was found at restart~10, iteration~2{,}807{,}777.
Because the search uses an unseeded pseudorandom generator
(\texttt{Math.random}), these iteration counts reflect a single
successful run and are not exactly reproducible. The published
witnesses are reproducibly verified by the Lean kernel regardless
of the search path that discovered them.

\section{Formal Verification}

Each witness is a static array in $\{0,\ldots,5\}^{m^3}$. We decode it
(Table~\ref{tab:permutations}) into three successor functions
$f_0, f_1, f_2 :
\texttt{Fin}\;m \times \texttt{Fin}\;m \times \texttt{Fin}\;m \to
\texttt{Fin}\;m \times \texttt{Fin}\;m \times \texttt{Fin}\;m$.
The decoding is formalized entirely within Lean; the successor
functions natively evaluate the lexicographic mapping
$(i,j,k) \mapsto im^2 + jm + k$ to index into the array.
Array access uses Lean's \texttt{getD} (total with a default
fallback), which is formally sound: any out-of-bounds access
would yield an invalid permutation and fail the subsequent
topological checks.

The specification uses Lean's \texttt{Prop} universe. For $m=6$, we
define the vertex type alias and the torus edge relation:
\begin{lstlisting}[language=ML,mathescape=true]
abbrev Torus6 := Fin 6 $\times$ Fin 6 $\times$ Fin 6

def isTorusEdge (u v : Torus6) : Prop :=
  v = (u.1 + 1, u.2.1, u.2.2) $\lor$
  v = (u.1, u.2.1 + 1, u.2.2) $\lor$
  v = (u.1, u.2.1, u.2.2 + 1)
\end{lstlisting}
We define the vertex coordinate space using Lean's core
\texttt{Fin\;m} type rather than the \texttt{ZMod\;m} ring provided by
Mathlib. This choice eliminates all external library dependencies,
keeping the trusted computing base (TCB) minimal---an approach also
utilized successfully in Morrison's formalization of the odd-$m$
cases~\cite{morrison2026}. Formally, our code verifies
the decomposition over this \texttt{Fin\;m} topology.
Metatheoretically, because Lean~4 defines addition on $\texttt{Fin}\;n$
as $(a + b) \bmod n$, this topology is structurally identical to the
standard mathematical object $(\mathbb{Z}/m\mathbb{Z})^3$. While
Mathlib provides \texttt{finEquivZMod} to formally witness this
equivalence, we omit the dependency to maintain a standalone kernel
verification.

Hamiltonicity is defined via finite iteration. A function $f$ on $m^3$
vertices is a Hamiltonian cycle iff its orbit from $(0,0,0)$ closes at
exactly $m^3$ steps and no shorter orbit exists. While an orbit of
maximal length on a finite set mathematically guarantees bijectivity,
we explicitly include injectivity in our Lean \texttt{Prop} definition
for definitional completeness.

Each proof consists of 15 sub-theorems verified by the \texttt{decide}
tactic:
\begin{lstlisting}[language=ML,mathescape=true]
theorem torus_m6_valid :
    IsHamiltonianDecomposition f0 f1 f2 :=
  $\langle$edges_f0, edges_f1, edges_f2,
   disj_01, disj_02, disj_12,
   $\langle$inj_f0, orbit_f0, no_short_f0$\rangle$,
   $\langle$inj_f1, orbit_f1, no_short_f1$\rangle$,
   $\langle$inj_f2, orbit_f2, no_short_f2$\rangle\rangle$
\end{lstlisting}

Lean formally verifies that the three successor functions are
mutually disjoint at every vertex (via \texttt{disj\_01},
\texttt{disj\_02}, \texttt{disj\_12}). Because the
\texttt{isTorusEdge} relation restricts the outgoing neighborhood
of any vertex to exactly three destinations, the complete partition
of the $3m^3$ arcs follows as a metatheoretical corollary of the
verified disjointness.

For $m=6$, the injectivity sub-theorems each check $216^2 = 46{,}656$ pairs,
and the no-sub-tour sub-theorems each verify 215 orbit prefixes of length
up to~215. We set \texttt{maxHeartbeats} to $4 \times 10^8$ to accommodate
the larger evaluation. Lean's \texttt{decide} tactic handles the full
verification without boolean reflection or \texttt{native\_decide}.
The $m=6$ proof compiles in 57 seconds (wall time) on an Apple~M4
processor, with zero unproven subgoals (\texttt{sorry}) and relying
only on Lean's standard axioms.

\section{Conclusion}

We have given machine-checked proofs that $G_4$ and $G_6$ each admit a
decomposition into three directed Hamiltonian cycles. Together with
Knuth's construction for odd $m \ge 3$~\cite{knuth2026} and the
GPT-5.4~Pro construction for even $m \ge 8$, this completes the
categorization for all $m \ge 3$.
(The degenerate cases $m=1$ and $m=2$ have fewer than $3m^3$ distinct arcs
and do not admit decompositions of this form.)

Source code, witnesses, and Lean~4 proofs are available at
\url{https://github.com/bneb/perqed}.

\section*{Acknowledgments}

We thank Donald Knuth for posing the problem and for his detailed
verification of the odd-$m$ construction~\cite{knuth2026}, Kim Morrison
for the contemporaneous Lean~4 formalization of the odd-$m$
case~\cite{morrison2026}, Ho Boon Suan for directing the GPT-5.4~Pro
discovery of the even-$m$ construction~\cite{hoboonsuan2026},
Keston Aquino-Michaels for further exploration of that
construction~\cite{aquino2026}, and the teams behind Claude Opus~4.6
(Anthropic) and GPT-5.4~Pro (OpenAI).

The TypeScript search engine, incremental energy evaluator, Lean~4 proof
architecture, and this manuscript were developed with assistance from
Claude (Anthropic) and Gemini (Google DeepMind), used as interactive
coding and writing assistants throughout the project.

\begin{thebibliography}{9}

\bibitem{knuth2026}
D.~E. Knuth, ``Claude's Cycles,'' Stanford Computer Science Department,
February 28, 2026 (revised March 6, 2026).
Available: \url{https://www-cs-faculty.stanford.edu/~knuth/papers/claude-cycles.pdf}

\bibitem{darijani2025}
I.~Darijani, B.~Miraftab, and D.~Witte~Morris, ``Arc-disjoint hamiltonian
paths in Cartesian products of directed cycles,'' \emph{Ars Mathematica
Contemporanea}, vol.~25, no.~2, P2.10, 2025.
DOI: \url{https://doi.org/10.26493/1855-3974.3047.c2d}

\bibitem{hoboonsuan2026}
H.~Boon Suan, ``even\_closed\_form.c'' and
``even\_closed\_form\_proof\_final.pdf,'' March 2026.
Available: \url{https://www-cs-faculty.stanford.edu/~knuth/even_closed_form.c}
and \url{https://www-cs-faculty.stanford.edu/~knuth/even_closed_form_proof_final.pdf}

\bibitem{aquino2026}
K.~Aquino-Michaels, ``Completing Claude's cycles: Multi-agent structured
exploration on an open combinatorial problem,'' No Way Labs, March 2026.
Available: \url{https://github.com/no-way-labs/residue}

\bibitem{morrison2026}
K.~Morrison, ``KnuthClaudeLean: Lean 4 formalization of Knuth's odd-$m$
torus decomposition,'' GitHub repository, March 4, 2026.
Available: \url{https://github.com/kim-em/KnuthClaudeLean}

\bibitem{kirkpatrick1983}
S.~Kirkpatrick, C.~D.~Gelatt, and M.~P.~Vecchi, ``Optimization by simulated
annealing,'' \emph{Science}, vol.~220, no.~4598, pp.~671--680, 1983.

\bibitem{marinari1992}
E.~Marinari and G.~Parisi, ``Simulated tempering: a new Monte Carlo scheme,''
\emph{Europhysics Letters}, vol.~19, no.~6, pp.~451--458, 1992.

\end{thebibliography}

\end{document}
